### 2020-02-13 introduction.tex
Features: safety

Over the past decade, there has been a significant increase in the number of
simulation tools made available for transient MSR analyses. They generally fall
into two categories: loosely coupled, separate, stand-alone codes for
neutronics and thermal-hydraulic, and multiphysics software for concurrently
solving neutron transport and thermal-hydraulics. Fiorina et al.
\cite{fiorina_modelling_2014} presented simulation results from two
models developed independently at \textit{Politechnico di Milano} and
\textit{Technical University of Delft} in the context of the Molten Salt Fast
Reactor (MSFR). The MSFR is a circulating fuel reactor concept under active
development in the H2020 SAMOFAR project. The fuel consists of a molten salt
mixture of thorium and uranium fluorides dissolved in lithium fluoride (LiF).
The former employed the commercial
multiphysics modeling software COMSOL \cite{comsol_comsol_nodate}, while the
latter coupled two in-house codes DALTON and HEAT for neutron transport and
CFD calculations respectively. Both models showed generally good agreement with
each other with a simplified, 2-D axisymmetric model for several transient
scenarios.

>>>>

Over the past decade, there has been a significant increase in the number of
simulation tools made available for transient MSR analyses. They generally fall
into two categories: loosely coupled, separate, stand-alone codes for
neutronics and thermal-hydraulic, and multiphysics software for concurrently
solving neutron transport and thermal-hydraulics. Fiorina et al.
\cite{fiorina_modelling_2014} presented simulation results from two
models developed independently at \textit{Politechnico di Milano} and
\textit{Technical University of Delft} in the context of the Molten Salt Fast
Reactor (MSFR). The MSFR is a circulating fuel reactor concept under active
development in the H2020 SAMOFAR project. The fuel consists of a molten salt
mixture of thorium and uranium fluorides dissolved in lithium fluoride (LiF).
The former employed the commercial
multiphysics modeling software COMSOL \cite{comsol_ab_comsol_2018}, while the
latter coupled two in-house codes DALTON and HEAT for neutron transport and
CFD calculations respectively. Both models showed generally good agreement with
each other with a simplified, 2-D axisymmetric model for several transient
scenarios.

### 2020-02-13 msr.tex
Features: safety

\glspl{MSR} hold a number of safety advantages over current \glspl{LWR}. The
most immediate and significant advantage is the strong negative temperature
reactivity feedback of the fuel salt. This is due to greater
temperature-induced density changes in liquid fuel than solid fuel. Combined
with the Doppler broadening of resonance capture cross sections present in
both fuel forms, we expect to observe a smaller temperature increase following
an unprotected reactivity insertion. The overall temperature reactivity
coefficient varies widely between different \gls{MSR} designs due to other
structures, such as moderators and reflectors, present in the core. In
particular, graphite moderators tend to have slightly positive temperature
reactivity coefficients. This was notably observed in the \gls{MSBR} concept,
but the total temperature reactivity coefficient was still relatively large
and negative.

Continuous online fuel reprocessing allows operators to maintain low excess
reactivity inventories in the core as additional fuel can be added on an ad
hoc basis. This factor, in addition to the strong negative temperature
reactivity coefficient, diminishes the likelihood and severity of unprotected
criticality accidents in \glspl{MSR}.

>>>>

\glspl{MSR} are generally deemed to be safer than \glspl{LWR} due to their
reliance on natural physical phenomena for passive safety.
The most immediate and significant safety characteristic is the strong
negative temperature reactivity feedback of the fuel salt. This is due to
greater temperature-induced volumetric expansion in liquid fuel than solid
fuel. Combined with the Doppler broadening of resonance capture cross sections
present in both fuel forms, we expect to observe a smaller temperature
increase following an unprotected reactivity insertion. The overall
temperature reactivity coefficient varies widely between different \gls{MSR}
designs due to other structures, such as moderators and reflectors, present in
the core. In particular, graphite moderators tend to have slightly positive
temperature reactivity coefficients. This was notably observed in the
\gls{MSBR} concept, but the total temperature reactivity coefficient was still
relatively large and negative \cite{rykhlevskii_modeling_2019}. This
phenomenon provides a large degree of control and stability as it is always
present in an \gls{MSR} regardless whether the operating conditions.

Continuous online fuel reprocessing allows operators to maintain low excess
reactivity inventories in the core as additional fuel can be added on an ad
hoc basis. Reprocessing and gas sparging systems help reduce fissile
requirements by continuously removing neutron poisons. These factors, in
addition to the strong negative temperature
reactivity coefficient, diminishes the likelihood and severity of unprotected
criticality accidents in \glspl{MSR} \cite{elsheikh_safety_2013}. In the
unlikely situation where an \gls{MSR} encounters a severe runaway reaction,
we can rely on another passive safety feature, freeze plugs. A freeze plug is
a plug filled with solidified salt at the bottom of the core
\cite{elsheikh_safety_2013}. It is actively
cooled by fans or other cooling systems to keep it just below the freezing
point of the salt. When temperatures in the core exceed a certain threshold
during a dangerous transient, the freeze plug melts and the molten salt in the
core drains into a containment tank designed to keep the salt in a
sub-critical configuration. This is especially easy to achieve for
thermal-spectrum \glspl{MSR} as the absence of moderators in the containment
tank would automatically bring the multiplication factor down below unity
\cite{elsheikh_safety_2013}. \glspl{MSR} also typically have large
margin-to-boiling under nominal operating conditions so that fuel salt boiling
does not occur. Furthermore, the reactor vessel is consequently subject to
much lower stresses as \glspl{MSR} operate at near-atmospheric pressure
levels. Thus, the probability of pipe ruptures due to high pressure is low.

\glspl{MSR} may be more resilient to pump failure accidents as natural
circulation could passively sustain enough heat transfer through the heat
exchanger to remove any remaining fission and decay heat after the failure. If
natural circulation proves insufficient, the aforementioned freeze plug can
drain the salt out of the core. Decay heat in \glspl{MSR} with online
reprocessing isexpected to be lower than that in \glspl{LWR} due to the
continuous removal of \glspl{FP}. For example, the decay heat in an \gls{MSFR}
after reaching equilibrium salt composition is approximately 3.5 \% compared
to 6 \% in \glspl{LWR} \cite{brovchenko_design-related_2013}.

However, there are also some safety disadvantages associated with \glspl{MSR}.
Firstly, \glspl{MSR} have smaller fractions of \gls{DNP} in the active core
region as some of them decay in the external loop regions. This complicates
reactor control and may result in faster transients due to the decrease in
average neutron lifetime. Secondly, structural materials in the core must be
resistant to salt corrosion under high temperatures and neutron irradiation
over the lifetime of the reactor. To a lesser extent, pipes should also be
corrosion resistant to prevent pipe ruptures. Lastly, fuel salt overcooling is
a potentially dangerous accident scenario as the strong temperature
coefficient can effectively cause a large reactivity insertion. Overcooling
may also cause salt to freeze and restrict flow, thereby causing a complete
loss of flow accident.

\subsection{Sustainability}
